---
title: "test"
author: "Min Hein Htike"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Prepare Data
```{r}
data(crossbill)
umf <- unmarkedFrameOccu(
    y=as.matrix(crossbill[,c("det991", "det992", "det993")]),
    siteCovs=crossbill[,c("ele", "forest")],
    obsCovs=list(date=crossbill[,c("date991", "date992", "date993")]))
sc <- scale(siteCovs(umf))
siteCovs(umf) <- sc
head(umf)
```
Notice that the site covariates, elevation and forest, were standardized using the scale function. Standardization isn’t always necessary, but it can make it easier to find the maximum likelihood estimates. When standardizing covariates and then making predictions, it is important to retain the original sample mean and standard deviation. The reason for this is explained below.

Fitting a model is now straight-forward. In many cases, we would fit several models corresponding to competing hypotheses, but for simplicity, we stick with this single model.

```{r}
(fm.occu <- occu(~date ~ele + I(ele^2) + forest, umf))
```

Now that we have our fitted model, we can use it to predict occurrence probability at each pixel in the Swiss landscape. The Switzerland dataset contains country-wide data. There are many ways to display it—here is an example of mapping elevation using the levelplot function in the lattice package (Sarkar 2008).

```{r}
library(lattice)
data(Switzerland)
print(levelplot(elevation ~ x + y, Switzerland, aspect="iso",
                xlab="Easting (m)", ylab="Northing (m)",
                col.regions=terrain.colors(100)))
```

The raster package (Jacob van Etten 2012) provides another alternative. Here we create two raster objects and specify the coordinate system.
```{r}
library(raster)
```

```{r}
elevation <- rasterFromXYZ(Switzerland[,c("x","y","elevation")],
    crs="+proj=somerc +lat_0=46.95240555555556 +lon_0=7.439583333333333 +k_0=1 +x_0=600000 +y_0=200000 +ellps=bessel +towgs84=674.374,15.056,405.346,0,0,0,0 +units=m +no_defs")

forest <- rasterFromXYZ(Switzerland[,c("x","y","forest")],
    crs="+proj=somerc +lat_0=46.95240555555556 +lon_0=7.439583333333333 +k_0=1 +x_0=600000 +y_0=200000 +ellps=bessel +towgs84=674.374,15.056,405.346,0,0,0,0 +units=m +no_defs")
```



Since we standardized the covariates during the model fitting process, we need to transform the country-wide data using the same values. Note, we don’t want to use the mean and SD of the rasters themselves, we want to use the mean and SD of the original covariates used to fit the models, which are stored as attributes of the sc object. The following commands display the original means and SDs and then transform the rasters and join them in a raster “stack”.

```{r}
attr(sc, "scaled:center")
```
```{r}
attr(sc, "scaled:scale")
```

```{r}
ele.s <- (elevation-1189)/640
forest.s <- (forest-34.7)/27.7
ef <- stack(ele.s, forest.s)
names(ef) <- c("ele", "forest")
plot(ef, col=terrain.colors(100))
```

It is important to assign names that exactly match the covariate names used to fit the model. This is required by the predict function as demonstrated later. The predict function is useful for computing spatially-referenced model predictions, standard errors, and confidence intervals, but it is computationally demanding when there are many pixels in the raster. Thus, if measures of uncertainty are not required, the following code can be used to quickly produce the species distribution map shown in Fig. 3.
```{r}
(beta <- coef(fm.occu, type="state"))
```

```{r}
logit.psi <- beta[1] + beta[2]*ele.s + beta[3]*ele.s^2 + beta[4]*forest.s
psi <- exp(logit.psi) / (1 + exp(logit.psi))
print(spplot(psi, col.regions=terrain.colors(100)))
```

Figure 3. A species distribution map for the European crossbill in Switzerland. The colors represent occurrence probability.

As of version 0.9-6, the predict method in unmarked can make predictions using an object of class RasterStack from the package. As mentioned previously, the rasters must be named, perhaps by using the names(someraster) <- somename method. The object returned by predict is another raster stack with rasters for the expected values of the parameter of interest, the standard errors, and the upper and lower confidence intervals. The following example is very slow because there are many of pixels in the raster. The resulting map is shown in Fig. 4.

```{r}
E.psi <- predict(fm.occu, type="state", newdata=ef)
plot(E.psi, axes=FALSE, col=terrain.colors(100))
```

Figure 4. Expected occurrence probability along with standard errors and the limits of the asymptotic 95% confidence interval.

Users should be cautious when predicting from models that have categorical predictor variables, i.e. factors. The package does not have advanced methods for handling factors, and thus it is not easy to automatically create dummy variables from them as can typically be done using model.matrix. The safest option is to create the dummy variables manually before fitting the models, and to use the same variables as rasters for prediction.

A more important consideration when creating species distribution maps based upon occurrence probability is that of spatial scale. Occurrence probability will typically depend upon the area of the “site” in question. Thus, in our crossbill example, it would not be appropriate to use our model to predict occurrence probability for 10-km2
 pixels since the surveys were done in 1-km2
 quadrats. In some cases it might be possible to directly model the effect of site area on occurrence probability, in which case the effect could be accounted for in the predictions.
 

## Interpreting relationship with habitat covariates

Fitted models can be used to predict expected outcomes when given new data. For example, one could ask “how many crossbills would you expect to find in a quadrat with 50% forest cover?” Prediction also offers a way of presenting the results of an analysis. We illustrate by plotting the predictions of ψ
 and p
 over the range of covariate values studied. Note that because we standardized date, we need to transform it back to its original scale after obtaining predictions on the standardized scale.
 
```{r}
nd <- data.frame (scale(crossbill[,c("ele", "forest")]))
E.psi <- predict(fm.occu, type="state", newdata=nd, appendData=TRUE)

predicted_df <- data.frame(predicted = E.psi$Predicted,lower = E.psi$lower, upper= E.psi$upper, (crossbill[,c("ele", "forest")]))

plot(predicted_df$ele, predicted_df$predicted)
plot(predicted_df$forest, predicted_df$predicted)
```
Let's make it beautiful
```{r}
p1 <- ggplot(predicted_df, aes(ele, predicted)) +
  geom_point(alpha = 0.3, colour = "red")+
  geom_smooth(span = 3)+
  ggtitle("Species' Prefered Elevation") +
  ylim(0, 1)+
  xlab("Elevation (m)")+
  ylab("Occupancy (pi)")

p2 <- ggplot(predicted_df, aes(forest, predicted)) +
  geom_point(alpha = 0.3, colour = "red")+
  geom_smooth(span = 1)+
  ggtitle("Species' Prefered Forest Cover") +
  ylim(0, 1)+
  xlab("Percentage of Forest cover(%)")+
  ylab("Occupancy (pi)")

grid.arrange(p1, p2)
```

 


